---
title: "Lab-5"
author: "Alain Perez"
date: "3/2/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(babynames)
library(tidyverse)
data(babynames)
glimpse(babynames)
```
We will now extract the "Bella" names.

```{r}
babynames%>%
filter(name =="Bella" & sex =="F") %>% head()

```

next we will display the whole babynames dataset retaining the varaibles year, name, and n.

```{r}
babynames%>% 
select(year, name, n)%>% head()

```

Now we will try chaining select and filter in order to get the smae list but for bellas only.

```{r}

Bellas<-babynames%>%
  filter(name =="Bella" & sex =="F") %>%
  select(year, name, sex, n)
head(Bellas)


```
Now we will seperate names into two groups a "popular" group and a "non-popular" group.

```{r}
modified_babynames<-babynames %>%
  mutate(popular = prop>=.01)
  head(modified_babynames)
```
Renaming "popular" to "is popular".

```{r}
modified_babynames<-modified_babynames%>%
  rename(is_popular = popular)
  head(modified_babynames)

```

##Excercise 1##

```{r}
PopularBabynames<-modified_babynames%>%
  filter(is_popular == TRUE)%>%
  select(-is_popular)
  head(PopularBabynames)


```

Now we can arrange the cases in descending order by popularity.

```{r}
babynames%>%
arrange(desc(prop))%>%
head()
```

##Excerise 2##

The prop variable is telling us what proportion of babies of a given sex and year had the given name. So to be first on the list means that the name in question was given to the most babies in a particuar year of a particular sex.

##Excercise 3##
```{r}
babynames%>%
  filter(year =="1997")%>%
  arrange(desc(prop))%>%
    head()

```
The most popular name in my birth year, 1997, was unfortunately not Alain. It was Michael.

##Excercise 4##
```{r}
babynames%>%
  filter(name == "Alain")%>%
  arrange(desc(prop))%>%
  head()
```
In 1987 the name "Alain" was most popular.

##Excercise 5##


```{r}
babynames%>%
  filter(name =="Alain")%>%
  arrange(desc(n))%>%
  head()
  
```

The year in which the greatest number of babies were named "Alain" is 1988 a full 91 babies were named "Alain" that year. This could be different from the proportion because in another year there couldve been a drastic decrease in births but the same number of Alains would have been borm resulting in a higher proportion but not an increase in popularity.

##Excercise 6##
I would attempt this question by initializing two varaibles one called male_counter and another called female_counter. They would count how many babies each year were named a given name and count how many of those were either males or females. Then I would have a varaible called ratio that equaled male_counter/female_counter and arrange the data set by decreasing order of ratios.

To prevent confusion we are going to rename the n varaible the num_births varaible.

```{r}
modified_babynames<-modified_babynames %>%
  rename(num_births = n)
  head(modified_babynames)

```

Finding out when Bella peaked

```{r}
Bellas<-modified_babynames%>%
  filter(name == "Bella" & sex == "F")
Bellas%>%
  summarize(num_rows =n(), most_bellas= max(num_births))


```

```{r}
Bellas %>%
  summarize(
    num_rows = n(),
    peak_number = max(num_births),
    peak_year = year[which.max(num_births)]
    
  )
```
##Excercise 7##
The num_rows variable tells us how many observations there are in the dataset so how many bellas that are female that are in the dataset.

##Excercise 8##
```{r}
birthyear<-filter(babynames, year =="1997")
birthyear%>%
summarize(num_rows=n(),most_pop=(max(prop)), name[which.max(prop)])

```

Using the summarize command the most popular name in 1997 was Michael.

```{r}
alain<-filter(babynames,name=="Alain")
alain%>%
  summarize(num_rows=n(),pop=max(prop), year[which.max(prop)])
```
As confirmed above the year where the name "Alain"" was the most popular was 1987

##Excercise 9##

```{r}
modified_babynames%>%
filter(name == "Colin" & sex == "M")%>%
  arrange(desc(num_births))%>%
    select(year, num_births)%>%
      head()
```